---
title: "Mini data analysis"
author: "Mathias Delhaye"
output: github_document
---

# Introduction
After installing the *datateachr* package in the R terminal, load the packages *datateachr* and *tidyverse* for data exploration 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(datateachr)
library(tidyverse)
```

# Task1: choosing my favorite dataset

**1. Choosing 4 data sets based on their descrition**

1: *steam_games*

2: *cancer_sample*

3: *vancouver_trees*

4: *apt_buildings*

**2. Exploring the data sets**

Let's explore the datasets

```{r}
dim(steam_games) #look at the dimension
glimpse(steam_games) #having a glimpse of what is in the data set
class(steam_games) #knowing what kind of data set it is
```

For this data set, the fact that there is the price as well as chain of characters variables is interesting for data exploring, and it is a tibble, making it easy to analyze

```{r}
dim(cancer_sample) #look at the dimension
glimpse(cancer_sample) #having a glimpse of what is in the data set
class(cancer_sample) #knowing what kind of data set it is
```

This data set is interesting, as all of the variables are numerical, allowing a lot of correlation, the dimensions are not that big and it is a tibble

```{r}
dim(vancouver_trees) #look at the dimension
glimpse(vancouver_trees) #having a glimpse of what is in the data set
class(vancouver_trees) #knowing what kind of data set it is
```

For this data set, several variables are numeric which is good to do analysis, it is a tibble and it has the neighborhood name

```{r}
dim(apt_buildings) #look at the dimension
glimpse(apt_buildings) #having a glimpse of what is in the data set
class(apt_buildings) #knowing what kind of data set it is
```

This data set is also a tibble, with some numerical variable that can be exploit and it is not that big. 

**3. Narrowing down to 2**

Based on the exploration before, the 2 data sets that seem the most promising are *cancer_samples* and *steam_games*

* vancouver_trees: there are some numerical variables as well as chains of characters, which increase the number analysis possible, not just correlation between numerical variable
* steam_games: this data set is interesting explore some relation between prices and genre of the game, in order to use not just numerical variable, but also chains of characters. And I love video games x)

**4. Final decision**

I am going to choose the *vancouver_trees* data set, as it presents variables that are numerical and chains of character that are short and simple to analyze, contrary to *steam-games*, that presents few numerical variable and the **chr** variables are very complex. 
For the *vancouver_trees*, I am going to explore the relation between physiological data of the trees (*diameter* and *height_range_id*) and other variables, such as the location or the type of tree. 

# Task2 exploring the chosen dataset

Let's dive in the *vancouver_trees* data set to see what is interesting to study! 

* Exercise 1

````{r}
vancouver_trees%>%
  filter(diameter < 60)%>% #Filtering to have a better view as most of the diameter are below 60
  ggplot(aes(diameter, height_range_id)) + #plotting the distribution of the height depending of the diameter
  geom_point(alpha = 0.1) #making the points a bit transparent to have a better view
````

I chose this exercise because I wanted to see the distribution of the height depending on the diameter, as it is 2 numerical variables from the data set related to the physiology of the trees

* Exercise 7

```{r}
vancouver_trees%>%
  group_by(neighbourhood_name)%>%
  summarise(n = n())%>% # I want to store the number of trees for each neighborhood
  ggplot(aes(neighbourhood_name,n)) + #plotting the number of trees per neighborhood
  geom_col() +
  coord_flip() # I had this line after realizing it was better to visualize
```

I chose this exercise because I found it interesting to determine the number of trees per neighborhood, it's a relation that can be useful in the future. I could have use the streets or the blocks, but for me it makes more sense to look at the neighborhoods because there is certainly an inner tree politic per neighborhood, and not per street

* Exercise 5 

```{r}
vancouver_trees%>%
  mutate(year_planted = format(date_planted, "%Y")) %>% #creating a new column with just the year when each tree was planted
  filter(year_planted<1996) #filtering to just take the trees planted before 1996
```

I chose this exercise because there is a lot of dates, and I wanted to filter by a year to restrain my analysis. Tt could be interesting to just look at some years, for example here before 1996. 

By the way here, we can see that our tibble is 10,000 rows while the original one is more than 100,000, meaning that it seems that more trees were planted from 1996

*Exercise 8

```{r}
vancouver_trees%>%
  filter(diameter<50)%>% #filtering because I noticed most of the trees have a diameter below 50, and with this filter the visualization is better
  ggplot(aes(x = diameter)) + 
    geom_density(aes(fill = root_barrier), alpha = 0.3)
```
For this last one, I wanted to check the relation between the diameter of trees and if there were root barrier for the tree or not, which could affect the growth of the tree.

# Task3: research questions

* Question 1

What is the evolution of trees plantation with the years? Indeed, by filtering with the year, I notice that I decreased the number of rows by 10, which may suggest a potential increase in the number of trees planted. This question can be extended to each neighborhood as variability in the number of trees per neighborhood was found

* Question 2

With the variability in the number of trees per neighborhood, can we also find some prevalence in some species for each neighborhood?

* Question 3

Is there a trend in the type of trees that are planted over the years? 

* Question 4

Is there an evolution of the physiological properties of the trees (diameter and height), depending on the year? Are these properties related to any other variables of the tibble?


# Comments

The questions I figured out seem a bit not enough for me, I have the feeling they are too simple, not deep enought for this analysis.